---
title: "Length Weight Relationship Evaluations"
author: "Adam Kemberling"
date: "Updated on: `r Sys.Date()`"
format: 
  html:
    toc: true
    self-contained: true
    code-fold: true
execute:
  echo: true
  fig-width: 8
  fig-height: 7
  message: false
  warning: false
editor: source
---



```{r}
# Libraries
library(tidyverse)
library(gmRi)
library(rnaturalearth)
library(scales)
library(sf)
library(gt)
library(patchwork)
library(lmerTest)
library(emmeans)
library(merTools)
library(tidyquant)
library(ggeffects)
library(performance)
library(gtsummary)
library(gt)
library(sizeSpectra)
library(ggdist)
library(pander)
library(ggh4x)
conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")
conflicted::conflicts_prefer(lmerTest::lmer)

# Processing functions
source(here::here("Code/R/Processing_Functions.R"))

# ggplot theme
theme_set(
  theme_gmri(
    rect = element_rect(fill = "white", color = NA), 
    strip.text.y = element_text(angle  = 0),
    axis.text.x = element_text(size = 8),
    legend.position = "bottom"))



# vectors for factor levels
area_levels <- c("GoM", "GB", "SNE", "MAB")
area_levels_long <- c("Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight")
area_levels_all <- c("Northeast Shelf", area_levels)
area_levels_long_all <- c("Northeast Shelf", area_levels_long)

# table to join for swapping shorthand for long-hand names
area_df <- data.frame(
  area = c("Scotian Shelf", "Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight", "Northeast Shelf"),
  survey_area = c("SS", "GoM", "GB", "SNE", "MAB", "Northeast Shelf"),
  area_titles = c("Scotian Shelf", "Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight", "Northeast Shelf"))


# Degree symbol
deg_c <- "\u00b0C"


```

# What are we checking?

Is `numlen_adj` properly balancing out with abundnace?

Is the estimated biomass from length-weight keys similar to the bulk biomass from the survey? 

Are there specific species biases, or seasonal issues?

What are the body-size percentiles for length and weight?

```{r}
#| label: load input data usecd to estimate spectra

# Data used for Wigley estimates
wigley_in <- read_csv(here::here("Data/processed/wigley_species_trawl_data.csv"))
finfish_in <- read_csv(here::here("Data/processed/finfish_trawl_data.csv"))
```



### Numlen Adjusted

Lengths are measured for all individuals typically. The data records are such that there is a row for each unique species-sex-length record and when there are multiple individuals in the sample with the same length measurement this is recorded in the `numlen` field. In some situations this is not possible, and lengths are taken from a subset of the total sample for each species. There was also a survey methodology change in 2008, and with this survey change there were adjustments applied to the `abundance` totals for each species. In both of these cases `numlen` values remain unaffected, and only capture the abundance of individuals at each length measured. To achieve consistency between `abundance` (which is the total of individuals by species) and the sum() of `numlen` we apply a proportional adjustment of `numlen`.

If properly done, `sum(numlen_adj)` should equal `abundance` for each tow.

```{r}

# Station Totals
wigley_station_abunds <- wigley_in %>% distinct(est_towdate, est_year, season, station, comname, catchsex, abundance, biomass_kg)
finfish_station_abunds <- finfish_in %>% distinct(est_towdate, est_year, season, station, comname, catchsex, abundance, biomass_kg) 

# Are sexes weighed separately for bulk abundance/biomass


# bulk station totals
wigley_station_totals <- wigley_station_abunds %>% 
  group_by(est_year, season) %>% 
  summarise(
    abundance = sum(abundance, na.rm = T),
    biomass_kg = sum(biomass_kg, na.rm = T),
    .groups = "drop")

# Average species per tow
finfish_station_abunds %>% 
  group_by(est_towdate, station) %>% 
  summarise(n_species = mean(n_distinct(comname)),
            .groups = "drop") %>% 
  summarise(avg_species = mean(n_species))

# Cod average number/weight per tow
wigley_station_abunds %>% 
  filter(comname == "atlantic cod") %>% 
  summarise(
    mean_abund = mean(abundance, na.rm = T),
    mean_biom = mean(biomass_kg, na.rm = T))



# Numlen based totals
wigley_numlen_totals <- wigley_in %>% group_by(est_towdate, est_year, season, station, comname, catchsex, abundance, biomass_kg) %>% 
  summarise(
    numlen = sum(numlen, na.rm = T),
    numlen_adj = sum(numlen_adj, na.rm = T),
    sum_weight_kg= sum(sum_weight_kg, na.rm = T),
    .groups = "drop") %>% 
  group_by(est_year, season) %>% 
  summarise(
    numlen = sum(numlen, na.rm = T),
    numlen_adj = sum(numlen_adj, na.rm = T),
    sum_weight_kg= sum(sum_weight_kg, na.rm = T),
    .groups = "drop")
  


ggplot() +
  geom_line(
    data = wigley_station_totals,
    aes(est_year, abundance, color = season, linetype = "abundance column")) +
    geom_line(
    data = wigley_numlen_totals,
    aes(est_year, numlen_adj, color = season, linetype = "numlen_adj column"),
    linewidth = 1) +
    facet_wrap(~season, nrow = 2) +
  labs(title = "Abundance & numlen_adj - same")
```


### Biomass and Estimated Biomass

```{r}

ggplot() +
  geom_line(
    data = wigley_station_totals,
    aes(est_year, biomass_kg, color = season, linetype = "biomass column")) +
    geom_line(
    data = wigley_numlen_totals,
    aes(est_year, sum_weight_kg, color = season, linetype = "estimated length-weight biomass"),
    linewidth = 1) +
    facet_wrap(~season, nrow = 2) +
  labs(title = "Biomass and Estimated L-W Biomass - close")
```



### Season & Species Fits


```{r}

hist(wigley_in$ind_weight_kg, breaks = 100)





# Get the total weights for each station, then do it each year for every species
mismatch_ranks <- function(x){
  
  # Sum across body sizes to get length weight totals
  # we want to lump together across sex because its unclear if biomass column is consistently sex specific
  species_summaries <- x %>% 
    group_by(est_towdate, est_year, comname, id, season) %>% 
    summarise(
      number_caught = sum(numlen_adj, na.rm = T),
      length_weight_biomass = sum(sum_weight_kg, na.rm = T),
      .groups = "drop") 
  
  # sum across distinct abundance and biomass totals to get tow summaries
  # use distinct abundance and biomass to catch if there is a sex distinction,
  # but still total to match above
  tow_totals <- x %>% 
    distinct(est_towdate, est_year, season, station, id,  comname, abundance, biomass_kg) %>% 
    group_by(est_towdate, est_year, comname, id, season) %>% 
    summarise(
      ship_measured_biomass = sum(biomass_kg, na.rm = T),
      .groups = "drop") 
  
  
    # join the two sources and get difference
   species_summaries <- species_summaries %>% 
     left_join(tow_totals) %>% 
     mutate(lw_bias = ship_measured_biomass - length_weight_biomass) 
    
  # Format rankings to plot
  n_species <- length(unique(species_summaries$comname))
  ranks <- species_summaries %>% 
    mutate(
      comname = fct_reorder(.f = comname, .x = lw_bias, mean),
      species_rank = as.numeric(comname),
      extremes = ifelse(species_rank >= n_species - 20, "Bias Towards Under-Estimation", NA),
      extremes = ifelse(species_rank <= 20, "Bias Towards Over-Estimation", extremes)) 
  
  return(ranks)
  
}






# Plot flags top and bottom 15 species that don't match well with survdat$biomass
mismatch_plot <- function(x, outlierz = T){
  
  # All ranks in one panel
  all_ranks <- x %>% 
    ggplot(aes(lw_bias, species_rank)) +
      geom_boxplot(aes(group = species_rank), color = "gray60", outliers = outlierz) +
      geom_boxplot(data = filter(x, is.na(extremes) == FALSE),
                   aes(group = species_rank, color = extremes), outliers = outlierz) +
      geom_vline(xintercept = 0, size = 0.5, color = "black", linetype = 2) +
      scale_x_continuous(labels = comma_format()) +
      scale_color_gmri() +
      guides("color" = guide_legend(title = "")) +
      labs(x = "Weight Estimation Bias\nShipboard Weight - LW Biomass (kg)",
           y = "Rank") +
    theme(
      # legend.position = "inside",
      # legend.position.inside = c(0.2, 1),
      legend.position = "top")
  
  # Focus on top 5
  top_5 <- x %>% 
    filter(extremes == "Bias Towards Under-Estimation") %>% 
    ggplot(aes(lw_bias, comname)) +
      geom_boxplot(aes(group = species_rank), 
                   color = gmri_cols("lv orange"), 
                   outlier.colour = "gray80", outliers = outlierz) +
      scale_color_gmri() +
      scale_x_continuous(labels = comma_format()) +
      labs(x = "",
           y = "", 
           subtitle = "L-W biomass < 'survdat$Biomass'") +
    theme(title = element_text(hjust = 0.5))
  
  # Focus on bottom 5
  bot_5 <- x %>% 
    filter(extremes == "Bias Towards Over-Estimation") %>% 
    ggplot(aes(lw_bias, comname)) +
      geom_boxplot(aes(group = species_rank), 
                   color = gmri_cols("gmri blue"), 
                   outlier.colour = "gray80", 
                   outliers = outlierz) +
      scale_color_gmri() +
      scale_x_continuous(labels = comma_format()) +
      labs(x = "Weight Estimation Bias\nShipboard Weight - LW Biomass (kg)",
           y = "", 
           subtitle = "L-W mass > 'survdat$Biomass'") +
    theme(title = element_text(hjust = 0.5))
    
  
  
  # Patchwork plot
  plot_out <- all_ranks | (top_5/bot_5)
  return(plot_out)
}
```



```{r}
#| fig-height: 10
#| fig-width: 8


# Just Wigley
wigley_ranks <- wigley_in %>% 
  mismatch_ranks() 

# Wigley mismatch plot
wigley_ranks %>% 
  mismatch_plot(outlierz = F)


```

### When were outlier events for Dogfish

```{r}

df_outliers <- check_outliers(wigley_ranks %>% filter(comname == "spiny dogfish") %>% pull(lw_bias)) %>% as.data.frame()

# bind_cols(filter(wigley_ranks, comname == "spiny dogfish"), df_outliers) 

# How many are outliers?
sum(df_outliers$Outlier)/nrow(df_outliers)*100


```



### Body Size Percentile Exploration


```{r}

```

