---
title: "Spectra Story"
author: "Adam Kemberling"
date: "Updated on: `r Sys.Date()`"
format: 
  html:
    toc: true
    self-contained: true
    code-fold: true
execute:
  echo: true
  fig-width: 8
  fig-height: 7
  message: false
  warning: false
editor: source
---



```{r}
# Libraries
library(tidyverse)
library(gmRi)
library(rnaturalearth)
library(scales)
library(sf)
library(gt)
library(patchwork)
library(lmerTest)
library(emmeans)
library(merTools)
library(tidyquant)
library(ggeffects)
library(performance)
library(gtsummary)
library(gt)
library(sizeSpectra)
library(ggdist)
library(pander)
conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")
conflicted::conflicts_prefer(lmerTest::lmer)

# Processing functions
source(here::here("Code/R/Processing_Functions.R"))

# ggplot theme
theme_set(
  theme_gmri(
    rect = element_rect(fill = "white", color = NA), 
    strip.text.y = element_text(angle  = 0),
    axis.text.x = element_text(size = 8),
    legend.position = "bottom"))



# vectors for factor levels
area_levels <- c("GoM", "GB", "SNE", "MAB")
area_levels_long <- c("Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight")
area_levels_all <- c("Northeast Shelf", area_levels)
area_levels_long_all <- c("Northeast Shelf", area_levels_long)

# table to join for swapping shorthand for long-hand names
area_df <- data.frame(
  area = c("Scotian Shelf", "Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight", "All"),
  survey_area = c("SS", "GoM", "GB", "SNE", "MAB", "Northeast Shelf"),
  area_titles = c("Scotian Shelf", "Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight", "Northeast Shelf"))


# Degree symbol
deg_c <- "\u00b0C"



# Function to get the Predictions
# Drop effect fits that are non-significant  ###
get_preds_and_trendsignif <- function(mod_x){
  modx_preds <- as.data.frame(
    # Model predictions
    ggpredict(
      mod_x, 
      terms = ~ est_year * survey_area * season) ) %>% 
    mutate(
      survey_area = factor(group, levels = area_levels),
      season = factor(facet, levels = c("Spring", "Fall")))
  
    # Just survey area and year
    modx_emtrend <- emtrends(
      object = mod_x,
      specs =  ~ survey_area*season,
      var = "est_year") %>% 
      as_tibble() %>% 
      mutate(
        zero = 0,
        non_zero = if_else(between(zero, lower.CL, upper.CL), F, T))
  
    # Preds with signif
    modx_preds %>% left_join(select(modx_emtrend, survey_area, season, non_zero))
  
}

```

## Storycrafting - Northeast Shelf Spectra

>I think we need to step back a bit and find the main story again--which is about how size structure of the NES fish community has changed over time, whether that varies by region, and what covariates are associated with those changes.  
The [Aug draft you developed](https://docs.google.com/document/d/1bOIdQgFZbUavtBXPf570ATz9VjpjVX9Ki5hOqVw71CI/edit) really focused on [comparing regression models to consider drivers based on the fish community size spectrum](https://adamkemberling.github.io/Spectra_Northeast/Code/R/modeling/lm_models/lm_summary_materials.html) as viewed from the perspective of the Wigley species set vs. the full species set.  
However, I didn't see figures of those spectra in that draft (maybe there are somewhere else and I've lost track).  The recent slide deck really focuses on seasonal differences in the size spectra and influence of temperature.  - KMills

This document will be split into two sections (the second potentially moving to its own doc). 


# Part 1: Summary of Trends

For the first section the focus is on understanding changes in the size structure. This will be done using length and bodymass spectra, large and small fish indices, and looking at how body mass size quantiles have moved over time.

This section should address: Has the size distribution changed over time, where has it changed, and some nuance around what aspects about the distribution have shifted.


## Starting from Square Uno

> I'd like for us to start by rebuilding the story of how the size structure has changed over time and regional differences.  
As such, I would like for us to look at plots of the size spectrum slope for (1) the all-species length spectrum and (2) the Wigley species biomass spectrum from annual, seasonal and regional perspectives.  
I am thinking this would be two panel plots (all-spp length, Wigley-spp biomass) with 5 subpanels each (NES, GOM, GB, SNE, MAB) that are set up to show lines by seasonal and annual time steps, much like the plot on the right of slide 7 in your most recent slide deck...except with NES added and an annual line added.  Keeping the significant trend lines in is helpful.  This is the first step, so when you have that ready, please share it with the group.  - KMills



```{r}
#| label: load input data usecd to estimate spectra

# Data used for Wigley estimates
wigley_in <- read_csv(here::here("Data/processed/wigley_species_trawl_data.csv"))
finfish_in <- read_csv(here::here("Data/processed/finfish_trawl_data.csv"))
```

#### Aside: Biomass Fraction in Each Community

If we need to make a decision about which community we want to use, here is the different biomass totals for each:


```{r}
# glimpse(wigley_in)

# Load raw
# General tidying only, removal of strata outside our study area, Spring and Fall only
# (removes inshore and rarely/inconsistently sampled strata etc.)
# shrimps removed
trawl_basic <- read_csv(here::here("Data/processed/trawl_clean_all_data.csv"))

# Biomass totals
raw_totals <- trawl_basic %>% 
  distinct(est_towdate, cruise6, station, stratum, tow, comname,  biomass_kg) %>%   pull(biomass_kg) %>% 
  sum()

wig_biomass_total <- wigley_in %>% 
  distinct(est_towdate, cruise6, station, stratum, tow, comname,  biomass_kg) %>%   pull(biomass_kg) %>% 
  sum()

ffish_biomass_total <- finfish_in %>% 
  distinct(est_towdate, cruise6, station, stratum, tow, comname,  biomass_kg) %>%   pull(biomass_kg) %>% 
  sum()

# Make a table for display
tibble(
  "Community" = c("All (including crustaceans)", "All finfish", "Wigley Species Subset"),
  "Biomass Total" = c(raw_totals, ffish_biomass_total, wig_biomass_total)) %>% 
  mutate(`% of Total` = (`Biomass Total` / raw_totals)*100) %>% 
  gt()

```



## A. Size Spectra Trends

For the figures in part 1: non-zero trends over time have been highlighted in the appropriate plot panels with linear trends overlaid.



```{r}
#| label: all species vs wigley species length and mass spectrum

# Load the three regional datasets
# wigley_lenspectra <- read_csv(
#   here::here("Data/processed/wigley_species_length_spectra.csv"))  %>% 
#   mutate(survey_area = factor(survey_area, levels = area_levels))
wigley_bmspectra <- read_csv(
  here::here("Data/processed/wigley_species_bodymass_spectra.csv")) %>% 
  mutate(survey_area = factor(survey_area, levels = area_levels))
finfish_lenspectra <- read_csv(
  here::here("Data/processed/finfish_length_spectra.csv")) %>% 
  mutate(survey_area = factor(survey_area, levels = area_levels))

# Load Shelf
finfish_lenspectra_shelf <- read_csv(
  here::here("Data/processed/shelfwide_finfish_length_spectra.csv"))
wigley_bmspectra_shelf <- read_csv(
  here::here("Data/processed/shelfwide_wigley_species_bodymass_spectra.csv"))




# Join them together
wigley_bmspectra_all <- bind_rows(wigley_bmspectra, wigley_bmspectra_shelf) %>% 
  mutate(metric = "bodymass_spectra",
         community = "Wigley Subset",
         survey_area = factor(survey_area, levels = area_levels_all))
ffish_lenspectra_all <- bind_rows(finfish_lenspectra, finfish_lenspectra_shelf) %>% 
    mutate(metric = "length_spectra",
           community = "Finfish Community",
         survey_area = factor(survey_area, levels = area_levels_all))
# both metrics+communities
spectra_results <-  bind_rows(wigley_bmspectra_all, ffish_lenspectra_all) %>% 
  mutate(
    survey_area = fct_relevel(survey_area, area_levels_all),
    metric = fct_rev(metric))

```




::: panel-tabset

### Wigley Species - Bodymass Spectra

```{r}

# Model linear trends in time
bmspectra_mod_wig_all <- lmer(
  formula = b ~ est_year * survey_area * season + (1|est_year),
  data = wigley_bmspectra_all)

# Get the predictions and flag whether they are significant
bmspectra_trend_predictions <- get_preds_and_trendsignif(bmspectra_mod_wig_all) %>% 
  mutate(model_id = "bodymass_spectra-Wigley Subset") %>% 
  separate(model_id, into = c("metric", "community"), sep = "-") %>% 
  mutate(metric = fct_rev(metric),
         survey_area = factor(survey_area, levels = area_levels_all))





# Plot over observed data
# Contrast seasonal differences




# Make the plot
bmspectra_trend_p <- ggplot() +
   geom_ribbon(
    data = filter(bmspectra_trend_predictions, non_zero == TRUE),
    aes(x, ymin = conf.low, ymax = conf.high, fill = season),
    linewidth = 1, alpha = 0.35) +
  geom_point(
    data = wigley_bmspectra_all,
    aes(est_year, b, color = season),
    size = 0.8, alpha = 0.5) +
  geom_ma(
    data = wigley_bmspectra_all,
    aes(est_year, b, color = season, linetype = "5-Year Moving Average"), 
    n = 5, ma_fun = SMA) +
  geom_line(
    data = filter(bmspectra_trend_predictions, non_zero == TRUE),
    aes(x, predicted, color = season, linetype = "Significant Trend"),
    linewidth = 1, alpha = 1) +
  scale_fill_gmri() +
  scale_color_gmri() +
  facet_grid(survey_area ~ metric*community, scales = "free") +
  labs(
    x = "Year",
    y = "Exponent of Body Mass Distribution",
    linetype = "Linetype",
    color = "Season",
    fill = "Season")
 
# Show plot
bmspectra_trend_p


# # Summary tables
# bmspectra_mod_wig_all %>% 
#   gtsummary::tbl_regression()   %>% 
#   add_global_p(keep = T) %>% 
#   bold_p(t = 0.05)  %>%
#   bold_labels() %>%
#   italicize_levels() %>% 
#   as_gt() %>% 
#   tab_header(title = "Wigley Community - Mass Spectra Trends")

```



### Full Finfish Community - Length Spectra


```{r}

# Model linear trends in time
lenspectra_mod_all <- lmer(
  formula = b ~ est_year * survey_area * season + (1|est_year),
  data = ffish_lenspectra_all)

# Get the predictions and flag whether they are significant
lenspectra_trend_predictions <- get_preds_and_trendsignif(lenspectra_mod_all) %>% 
  mutate(model_id = "length_spectra-Finfish Community") %>% 
  separate(model_id, into = c("metric", "community"), sep = "-") %>% 
  mutate(metric = fct_rev(metric),
         survey_area = factor(survey_area, levels = area_levels_all))





# Plot over observed data
# Contrast seasonal differences




# Make the plot
lenspectra_trend_p <- ggplot() +
   geom_ribbon(
    data = filter(lenspectra_trend_predictions, non_zero == TRUE),
    aes(x, ymin = conf.low, ymax = conf.high, fill = season),
    linewidth = 1, alpha = 0.35) +
  geom_point(
    data = ffish_lenspectra_all,
    aes(est_year, b, color = season),
    size = 0.8, alpha = 0.5) +
  geom_ma(
    data = ffish_lenspectra_all,
    aes(est_year, b, color = season, linetype = "5-Year Moving Average"), 
    n = 5, ma_fun = SMA) +
  geom_line(
    data = filter(lenspectra_trend_predictions, non_zero == TRUE),
    aes(x, predicted, color = season, linetype = "Significant Trend"),
    linewidth = 1, alpha = 1) +
  scale_fill_gmri() +
  scale_color_gmri() +
  facet_grid(survey_area ~ metric*community, scales = "free") +
  labs(
    x = "Year",
    y = "Exponent of Body Length Distribution",
    linetype = "Linetype",
    color = "Season",
    fill = "Season")
 
# Show the plot
lenspectra_trend_p



# # Summary tables
# lenspectra_mod_all  %>% 
#   gtsummary::tbl_regression()   %>% 
#   add_global_p(keep = T) %>% 
#   bold_p(t = 0.05)  %>%
#   bold_labels() %>%
#   italicize_levels() %>% 
#   as_gt() %>% 
#   tab_header(title = "Full Finfish Community - Length Spectra Trends")
```

### Both Community Size Spectra


```{r}
#| label: plot all trends together


# Left side:
# All species  species length spectrum 
# color = season vs. annual
# facet_wrap(~survey_area)

# Right side:
# Wigley species biomass spectrum
# color = season vs. annual
# facet_wrap(~survey_area)


# Plot over observed data
# Contrast seasonal differences
(
  (lenspectra_trend_p + theme(strip.text.y = element_blank()) | 
  bmspectra_trend_p)
  ) + plot_layout(guides = "collect")
```


:::




## B. Median Length & Weight Trends


>I also think it would be valuable to build out the story of size change further, so am thinking that plots like those of average (or median) length (all species) and weight (wigley species) for NES and the sub-regions, like those on p. 12 of the attached file above (the very first draft from late 2022) would be useful.  - KMills  




```{r}
#| label: length changes


# Load the median weight/length data
wigley_size_df <- read_csv(here::here("Data/processed/wigley_species_size_summary.csv"))
finfish_size_df <- read_csv(here::here("Data/processed/finfish_species_length_summary.csv"))

# shelf-wide summary
finfish_lengths_shelf <- read_csv(here::here("Data/processed/shelfwide_finfish_species_length_summary.csv"))
wigley_sizes_shelf <- read_csv(here::here("Data/processed/shelfwide_wigley_species_size_summary.csv"))




# Join them
size_results <- bind_rows(
  list(
    "Finfish Community" = bind_rows(finfish_size_df, finfish_lengths_shelf),
    "Wigley Subset" = bind_rows(wigley_size_df, wigley_sizes_shelf)), 
  .id = "community"
) %>% 
  mutate(
    survey_area = factor(survey_area, levels = area_levels_all)
  )




# Get trends:
len_mod_ffish <- lmer(
  formula = med_len_cm ~ est_year * survey_area * season + (1|est_year),
  data = finfish_size_df)
# len_mod_wig <- lmer(
#   formula = med_len_cm ~ est_year * survey_area * season + (1|est_year),
#   data = wigley_size_df)
wt_mod_wig <- lmer(
  formula = med_wt_kg ~ est_year * survey_area * season + (1|est_year),
  data = drop_na(wigley_size_df, med_wt_kg))



# Get the predictions and flag whether they are significant
size_trend_predictions <- bind_rows(list(
  "med_len_cm-Finfish Community" = get_preds_and_trendsignif(len_mod_ffish),
  #"med_len_cm-Wigley Subset"     = get_preds_and_trendsignif(len_mod_wig),
  "med_wt_kg-Wigley Subset"   = get_preds_and_trendsignif(wt_mod_wig)
  ), .id = "model_id") %>% 
  separate(model_id, into = c("metric", "community"), sep = "-") %>% 
  mutate(metric = fct_rev(metric))






# Left side: 
# Median length - all species
# color = season vs. annual
# facet_wrap(~survey_area)
size_long <- size_results %>% 
  pivot_longer(cols = c(med_len_cm, med_wt_kg), 
               names_to = "metric", 
               values_to = "val")  %>% 
  mutate(survey_area = fct_relevel(survey_area, area_levels_all),
         season = fct_rev(season))


# just length for scales
len_plot <-  size_long %>% 
  filter(metric == "med_len_cm",
         community == "Finfish Community") %>% 
  drop_na(val) %>% 
  ggplot() +
  geom_ribbon(
    data = filter(size_trend_predictions, metric == "med_len_cm", non_zero == TRUE),
    aes(x, ymin = conf.low, ymax = conf.high, fill = season),
    linewidth = 1, alpha = 0.35) +
  geom_point(
    #data = filter(size_long, metric == "med_len_cm"),
    aes(est_year, val, color = season),
    size = 0.8, alpha = 0.5) +
  geom_ma(
    aes(est_year, val, color = season, linetype = "5-Year Moving Average"), 
    n = 5, ma_fun = SMA) +
  geom_line(
    data = filter(size_trend_predictions, metric == "med_len_cm", non_zero == TRUE),
    aes(x, predicted, color = season, linetype = "Significant Trend"),
    linewidth = 1, alpha = 1) +
  scale_fill_gmri() +
  scale_color_gmri() +
  facet_grid(survey_area ~ metric*community,
             scales = "free") +
  theme(strip.text.y = element_blank()) +
  labs(
    x = "Year",
    y = "Length (cm)",
    color = "Season",
    linetype = "Linetype",
    fill = "Season")

# Right: 
# Median weight - wigley species
# color = season vs. annual
# facet_wrap(~survey_area)

# weight plots
wt_plot <- size_long %>% 
  filter(metric == "med_wt_kg") %>% 
  drop_na(val) %>% 
  ggplot() +
  geom_ribbon(
    data = filter(size_trend_predictions, metric == "med_wt_kg", non_zero == TRUE),
    aes(x, ymin = conf.low, ymax = conf.high, fill = season),
    linewidth = 1, alpha = 0.35) +
  geom_point(
    aes(est_year, val, color = season),
    size = 0.8, alpha = 0.5) +
  geom_ma(
    aes(est_year, val, color = season, linetype = "5-Year Moving Average"), 
    n = 5, ma_fun = SMA) +
  geom_line(
    data = filter(size_trend_predictions, metric == "med_wt_kg", non_zero == TRUE),
    aes(x, predicted, color = season),
    linewidth = 1, alpha = 1) +
  scale_fill_gmri() +
  scale_color_gmri() +
  facet_grid(survey_area ~ metric*community,
             scales = "free")  +
  labs(
    x = "Year",
    y = "Weight (kg)",
    color = "Season",
    linetype = "Linetype",
    fill = "Season")

(len_plot | wt_plot) + plot_layout(guides = "collect") 
# (len_plot | wt_plot) + plot_layout(guides = "collect", widths = c(3,1.5)) & theme(legend.position = "bottom")

```

**Notes:**

There are a number of years when median weight surges in MAB. For these situations I think we likely could trace these down to exceptionally high catches of larger species like elasmobranchs.

We also don't see a shelf-wide decline in size so I might need to check what Friedland did or prepare for a confrontation.


## C. Small/Large Fish Trends

Data for part C. comes from the Wigly species subset to set percentiles based on individual bodymass.

The following figures frame "large fish" as individuals with bodymass greater than the 90th percentile for all years within that region. "Small fish" are considered smaller than the 10th percentile of the mass distribution.

The following table shows what those percentiles are for each region. 

**NOTE:**
Percentiles are determined using `DescTools::Quantile()` which uses `numlen_adj` to weight everything by abundance.


```{r}
#| label: weighted percentiles

# # Get 95th percentile as threshold
# DescTools::Quantile(
#   wigley_in$length_cm, 
#   weights = wigley_in$numlen_adj, 
#   probs = c(0.05, 0.1, 0.5, 0.90, .95)) %>% 
#   t() %>% 
#   as_tibble()


# Do them separately
area_size_quants <- bind_rows(
    mutate(wigley_in, survey_area = "Northeast Shelf"),
    wigley_in) %>% 
  split(.$survey_area) %>% 
  map_dfr(function(x,y){
    
    DescTools::Quantile(
      # x$length_cm, 
      x$ind_weight_kg,
      weights = x$numlen_adj, 
      probs = c(0.05, 0.1, 0.5, 0.90, .95)) %>% 
      t() %>% 
      as_tibble() 
  }, .id = "survey_area") %>% 
  mutate(survey_area = factor(survey_area, levels = area_levels_all)) %>% 
  arrange(survey_area)

# Show the table
area_size_quants %>% 
  mutate_if(is.numeric, round, 2) %>% 
  gt() %>% 
  tab_header(title = "Wigley Species - Body Mass Quantiles (kg)") %>% tab_spanner(label = "Percentile", columns = -survey_area) %>% 
  cols_label("survey_area" = "")

```


```{r}
#| label: assign large and small labels


##### Large Fish Index  ####


# Assign large/small
wigley_in <- wigley_in %>% 
  left_join(select(area_size_quants, survey_area, `10%`, `90%`)) %>%   
  mutate(
    small = ifelse(ind_weight_kg < `10%`, T, F),
    large = ifelse(ind_weight_kg > `90%`, T, F))



# Get total abundance
total_abundbio <- bind_rows(
    mutate(wigley_in, survey_area = "Northeast Shelf"),
    wigley_in) %>% 
  group_by(survey_area, est_year, season) %>% 
  summarise(
    total_abund = sum(numlen_adj),
    total_bio = sum(sum_weight_kg),
    .groups = "drop")
  
  
# Get large fish bio
lf_abundbio <- bind_rows(
    mutate(wigley_in, survey_area = "Northeast Shelf"),
    wigley_in) %>% 
  group_by(survey_area, est_year, season) %>% 
  summarise(
    small_abund = sum(numlen_adj * small),
    small_bio = sum(sum_weight_kg * small),
    large_abund = sum(numlen_adj * large),
    large_bio = sum(sum_weight_kg * large),
    .groups = "drop")

# Join and divide
lfi <- left_join(total_abundbio, lf_abundbio) %>% 
  mutate(
    LFI = large_abund / total_abund,
    SFI = small_abund / total_abund)  %>% 
  mutate(survey_area = factor(survey_area, levels = area_levels_all)) %>% 
  left_join(
    select(area_size_quants, survey_area, `10%`, `90%`) %>% 
      mutate(across(-survey_area, \(x) str_c(round(x, 2), " kg"))))
```

The following species are those which has representation in the large fish index:

```{r}
big_fishes <- wigley_in %>% 
  filter(large) %>% 
  # filter(comname == "acadian redfish") %>% glimpse()
  distinct(comname) %>% 
  arrange(comname) %>% 
  rename(`Common Name:` = comname) 


# Print it neatly
pander(big_fishes$`Common Name:`)

```

You might think: "Why is a 1kg acadian redfish a "large fish"?", and thats a consequence of abundance declining with size pretty dramatically.

What we may want to do is look at the distribution of all sizes, and not their frequency of occurence when setting percentiles.

::: panel-tabset

### Shifting 10th & 90th Percentiles

A 90th percentile fish is smaller than I feel it has been in the past, so I wanted to plot those quantiles over time. 

Indeed, the 90th & 95th percentile size for fishes in the GOM was nearly 2x what it is today.


```{r}
# Do them separately
yearly_size_quants <- bind_rows(
    mutate(wigley_in, survey_area = "Northeast Shelf"),
    wigley_in) %>% 
  unite(col = "group_var", survey_area, est_year, sep = "X", remove = T) %>% 
  split(.$group_var) %>% 
  map_dfr(function(x,y){
    
    DescTools::Quantile(
      # x$length_cm, 
      x$ind_weight_kg,
      weights = x$numlen_adj, 
      probs = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.90, .95)) %>% 
      t() %>% 
      as_tibble() 
  }, .id = "group_var") %>% 
  separate(group_var, into = c("survey_area", "est_year"), sep = "X") %>% 
  mutate(
    est_year = as.numeric(est_year),
    survey_area = factor(survey_area, levels = area_levels_all)) %>% 
  arrange(survey_area)


# Reshape to make filtering easier
quants_long <- yearly_size_quants %>% 
  #select(-c(`5%`,`50%`, `95%`)) %>% 
  pivot_longer(cols = -c(survey_area, est_year), names_to = "Percentile", values_to = "biomass_kg") 


# Separate axes
q10_p <- quants_long %>%
  filter(Percentile == "10%") %>% 
  ggplot(aes(est_year, biomass_kg)) +
  geom_point(size = 0.8, alpha = 0.5) +
  geom_ma(n = 5, ma_fun = SMA, aes(linetype = "5-Year Moving Average"), color = "black") +
  scale_color_gmri() +
  theme(
    strip.text.y = element_blank(),
    plot.margin = margin(0,0,0,0)) +
  facet_grid(survey_area~Percentile) +
  labs(y = "Bodymass (kg)", x = "Year", linetype = "")

# Separate axes
q90_p <- quants_long %>%
  filter(Percentile == "90%") %>% 
  ggplot(aes(est_year, biomass_kg)) +
  geom_point(size = 0.8, alpha = 0.5) +
  geom_ma(n = 5, ma_fun = SMA, aes(linetype = "5-Year Moving Average"), color = "black") +
  scale_color_gmri(reverse = T) +
  theme(
    strip.text.y = element_blank(),
    plot.margin = margin(0,0,0,0)) +
  facet_grid(survey_area~Percentile) +
  labs(y = "", x = "Year", linetype = "")


# Separate axes
q95_p <- quants_long %>%
  filter(Percentile == "95%") %>% 
  ggplot(aes(est_year, biomass_kg)) +
  geom_point(size = 0.8, alpha = 0.5) +
  geom_ma(n = 5, ma_fun = SMA, aes(linetype = "5-Year Moving Average"), color = "black") +
  scale_color_gmri(reverse = T) +
  theme(plot.margin = margin(0,0,0,0)) +
  facet_grid(survey_area~Percentile) +
  labs(y = "", x = "Year", linetype = "")

(q10_p | q90_p | q95_p) + plot_layout(guides = "collect", axes = "collect_x") +
  plot_annotation(title = "Regional Size Percentile Changes")


```

10th percentile weights seem stable. 90th percentile body-weights have fluctuated over time.


```{r}
#| label : percentiles as filled contours



# Map over years and areas?
quant_fillers <- yearly_size_quants %>% 
  unite(col = "year_area", sep = "X", est_year, survey_area, remove = F) %>% 
  split(.$year_area) %>% 
  map_dfr(function(x){
    tribble(
      ~"quant_range",   ~"x",            ~"y",       ~"xend",         ~"yend",
       "0-10%",   x[["est_year"]], 0,          x[["est_year"]], x[["10%"]],
       "10-20%",   x[["est_year"]], x[["10%"]], x[["est_year"]], x[["20%"]],
      #"0-20%",   x[["est_year"]], 0, x[["est_year"]], x[["20%"]],
      # Normal below
      "20-30%",   x[["est_year"]], x[["20%"]], x[["est_year"]], x[["30%"]],
      "30-40%",   x[["est_year"]], x[["30%"]], x[["est_year"]], x[["40%"]],
      "40-50%",   x[["est_year"]], x[["40%"]], x[["est_year"]], x[["50%"]],
      "50-60%",   x[["est_year"]], x[["50%"]], x[["est_year"]], x[["60%"]],
      "60-70%",   x[["est_year"]], x[["60%"]], x[["est_year"]], x[["70%"]],
      "70-80%",   x[["est_year"]], x[["70%"]], x[["est_year"]], x[["80%"]],
      "80-90%",   x[["est_year"]], x[["80%"]], x[["est_year"]], x[["90%"]]#,
      # Don't go to 100
      #"90-95%",  x[["est_year"]], x[["90%"]], x[["est_year"]], x[["95%"]]
      )

  }, .id = "year_area") %>% 
  separate(year_area, into = c("est_year", "survey_area"), sep = "X") %>% 
  mutate(
    quant_range = factor(quant_range, levels = c(
     "0-10%", "10-20%", 
     #"0-20%", 
     "20-30%", "30-40%", "40-50%", 
     "50-60%", "60-70%", "70-80%", "80-90%", "90-95%")),
    survey_area = factor(survey_area, levels = area_levels_all))
  


# Try that funky guy
ggplot(quants_long) +
  geom_ribbon(
    data = quant_fillers,
    aes(x = x, ymin = y*1000, ymax = yend*1000, fill = quant_range), 
    alpha = 1, color = "transparent", linewidth = 0.1) +
  geom_line(
    aes(est_year, biomass_kg*1000, group = Percentile), 
    linewidth = 0.2, alpha = 0.35, color = "white") +
  
  facet_grid(survey_area~., scales = "free") +
  guides(fill = guide_legend(nrow = 2, title.position = "top", title.hjust = 0.5)) +
  # This one looks good on true y values
  scale_fill_brewer(direction = 1) +
  #scale_y_continuous(labels = label_number(big.mark = ",")) +
  # # This one looks good on log10 y values
  # rcartocolor::scale_fill_carto_d(palette = "BluYl",direction = -1) +
  scale_x_continuous(expand = expansion(add = c(0,0))) +
  # theme_bw() + theme(
  #   legend.position = "bottom", 
  #   strip.background = element_rect(fill = "#00736D"), 
  #   strip.text = element_text(face = "bold", color = "white"),
  #   strip.text.y = element_text(angle = 0),
  #   plot.title = element_text(face = "bold")) +
  scale_y_log10() +
  labs(
    y = "Individual Bodymass (g)", 
    x = "Year", 
    fill = "Body Size Quantile Range",
    title = "Looks like an ice shelf")

```



### LFI & SMI

The axes on these are: $$\frac{\sum{individuals_{ijs}
> mass~threshold}}
{\sum{individuals_{ijs}}} * 100$$ for each: $year_i$, $region_j$, $season_s$.

```{r}
#| label: plot lfi as percent

# Plot them as loing-term average

# Plot the SFI
sfi_p <- lfi %>% 
  ggplot() +
  geom_point(
    aes(est_year, SFI, color = season),
    size = 0.8, alpha = 0.5) +
  geom_ma(
    aes(est_year, SFI, color = season, linetype = "5-Year Moving Average"), 
    n = 5, ma_fun = SMA) +
  scale_fill_gmri() +
  scale_color_gmri() +
  scale_y_continuous(labels = label_percent()) +
  theme(strip.text.y = element_blank()) +
  facet_grid(survey_area~., scales = "free") +
  labs(
    x = "Year",
    y = "Percent of Total Abundance",
    title = "Small Fish Index\n(< 10th Percentile Bodymass)",
    fill = "Season",
    color = "Season",
    linetype = "")


# Plot the LFI
lfi_p <- ggplot(lfi) +
  geom_point(
    aes(est_year, LFI, color = season),
    size = 0.8, alpha = 0.5) +
  geom_ma(
    aes(est_year, LFI, color = season, linetype = "5-Year Moving Average"), 
    n = 5, ma_fun = SMA) +
  scale_fill_gmri() +
  scale_color_gmri() +
  scale_y_continuous(labels = label_percent()) +
  facet_grid(survey_area~., scales = "free") +
  labs(
    x = "Year",
    y = "",
    title = "Large Fish Index\n(> 90th Percentile Bodymass)",
    fill = "Season",
    color = "Season",
    linetype = "")



# Plot together
(sfi_p | lfi_p) + plot_layout(guides = "collect")

```

### Large and Small Fish Abundance

This plots the abundance of individuals above/below the size thresholds as abundance.

```{r}
#| label: plot lfi in abundance

# Plot them as loing-term average

# Plot the SFI
sfi_p <- lfi %>% 
  ggplot() +
  geom_point(
    aes(est_year, small_abund, color = season),
    size = 0.8, alpha = 0.5) +
  geom_ma(
    aes(est_year, small_abund, color = season, linetype = "5-Year Moving Average"), 
    n = 5, ma_fun = SMA) +
  scale_fill_gmri() +
  scale_color_gmri() +
  scale_y_continuous(labels = label_number()) +
  theme(strip.text.y = element_blank()) +
  facet_grid(survey_area~., scales = "free") +
  labs(
    x = "Year",
    y = "Total Abundance",
    title = "Small Fish Index\n(< 10th Percentile Bodymass)",
    fill = "Season",
    color = "Season",
    linetype = "")


# Plot the LFI
lfi_p <- ggplot(lfi) +
  geom_point(
    aes(est_year, large_abund, color = season),
    size = 0.8, alpha = 0.5) +
  geom_ma(
    aes(est_year, large_abund, color = season, linetype = "5-Year Moving Average"), 
    n = 5, ma_fun = SMA) +
  scale_fill_gmri() +
  scale_color_gmri() +
  scale_y_continuous(labels = label_number()) +
  facet_grid(survey_area~., scales = "free") +
  labs(
    x = "Year",
    y = "",
    title = "Large Fish Index\n(> 90th Percentile Bodymass)",
    fill = "Season",
    color = "Season",
    linetype = "")




(sfi_p | lfi_p) + plot_layout(guides = "collect")

```

:::




# Summary of Trends


> Let's start with these pieces, and we can then discuss doing something with the extremes of the size distribution--changes in small vs. large individuals, where we pre-define size bins as you've done in slides 7 and 8, or perhaps using an approach similar to Lora's approach of defining the upper/lower percentiles of size. - KMillz




----

# Part 2: Relationships to External Factors

The aim of this section is to provide context and potentially attribution/correlations that may help explain trends/changes observed in the above section.


```{r}

# # Model Dataset for Wigley Species Subset
wigley_bmspectra_model_df <- read_csv(here::here("Data/model_ready/wigley_community_bmspectra_mod.csv"))

# Use one for plots
wigley_bmspectra_model_df <- wigley_bmspectra_model_df %>% 
  mutate(survey_area = case_when(
    survey_area == "GoM" ~ "Gulf of Maine",
    survey_area == "GB" ~ "Georges Bank",
    survey_area == "SNE" ~ "Southern New England",
    survey_area == "MAB" ~ "Mid-Atlantic Bight"),
  survey_area = factor(survey_area, levels = area_levels_long),
  season = fct_rev(season))

```


###  Bottom Temperature Changes

Bottom temperatures are from the DuPontavice and Saba combined bottom temperature product. This data product has been clipped to the study areas and seasonally averaged to match the survey sampling seasons.


```{r}



# Plot temperatures
temp1 <- wigley_bmspectra_model_df %>% 
  ggplot(aes(est_year, bot_temp, color = season)) +
  geom_point(size = 0.8, alpha = 0.5) +
  geom_ma(size = 1, n = 5, ma_fun = SMA, linetype = 1) +
  scale_color_gmri() + 
  facet_grid(survey_area~.) +
  scale_y_continuous(labels = label_number(suffix = "\u00b0C")) +
  labs(y = "Bottom Temperature", x = "Year", color = "5-Year Smooth")



temp1 
```

Bottom temperatures have increased in all four regions. The distribution of seasonal bottom temperatures are such that Gulf of Maine experiences a much smaller difference in seasonal temperature swings than the other regions.

### Bottom Temperature Distributions

Bottom temperatures are more seasonal in GB, SNE, and MAB. In the Gulf of Maine the difference in bottom temperatures


```{r}

# Plot the distribution as a raincloud plot
wigley_bmspectra_model_df %>% 
  ggplot(aes(fct_rev(survey_area), bot_temp, color = season, fill = season)) +
   ggdist::stat_halfeye(
     alpha = 0.4,
     adjust = .5, 
     width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
  ) + 
  geom_boxplot(
    fill = "transparent",
    width = .15, 
    outlier.shape = NA
  ) +
  ## add dot plots from {ggdist} package
  ggdist::stat_dots(
    ## orientation to the left
    side = "left", size = 1,
    ## move geom to the left
    justification = 1.12, 
    ## adjust grouping (binning) of observations 
    binwidth = .25) + 
  scale_color_gmri() + 
  scale_fill_gmri() + 
  scale_y_continuous(labels = label_number(suffix = "\u00b0C")) +
  coord_flip() +
  theme(legend.position = "none") +
  labs(y = "Bottom Temperature", x = "", color = "Season")
```

The overlap in bottom temperature distributions is very similar to the overlap in mass spectra exponents. Gulf of Maine spectra and bottom temperatures share a good region of overlap, and the regions that experience larger seasonal fluctuations in BT see larger seasonal fluctuations in their size distributions.

```{r}
#| label: comparison across regions


wigley_b_dist_plot <- wigley_bmspectra_model_df %>% 
  ggplot(aes(fct_rev(survey_area), b, color = season, fill = season)) +
  ggdist::stat_halfeye(
     alpha = 0.4,
     adjust = .5, 
     width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA) + 
  geom_boxplot(
    fill = "transparent",
    width = .15, 
    outlier.shape = NA) +
  ## add dot plots from {ggdist} package
  ggdist::stat_dots(
    ## orientation to the left
    side = "left", size = 1,
    ## move geom to the left
    justification = 1.12, 
    ## adjust grouping (binning) of observations 
    binwidth = .005) + 
  scale_color_gmri() + 
  scale_fill_gmri() + 
  coord_flip() +
  labs(
    y = "ISD Exponent", 
    x = "",
    title = "Individual Body Mass Distribution (1g - Max)")



# Plot next to bottom temperatures
wigley_b_dist_plot
```

This makes me lean towards belief that the size distribution is more responsive on shorter time scales to the ambient water temperatures. 

This to me also suggests that as a whole the community is less stationary and better able to respond to slower long-term trends in any individual region.


```{r}
wigley_bmspectra_model_df %>% 
  ggplot(aes(bot_temp, b)) +
  geom_point(aes(color = season, shape = survey_area)) +
  geom_smooth(method = "lm", color = "black") +
  scale_color_gmri() +
  scale_x_continuous(labels = label_number(suffix = deg_c)) +
  labs(title = "Broad Relationship to Ambient Bottom Temperature")
```


### Total Landings

The landings information is the newer GARFO data obtained from Andy Beet. It was delivered as containing only finfish, I have removed freshwater species and blue crabs. It should largely reflect marine finfish species at this point.


```{r}
wigley_bmspectra_model_df %>% 
  distinct(est_year, survey_area, total_weight_lb) %>% 
  ggplot(aes(est_year, total_weight_lb/1e6)) +
  geom_col(alpha = 0.5) +
  geom_ma(aes(color = "5-Year Smooth"), size = 1, n = 5, ma_fun = SMA, linetype = 1) +
  scale_color_manual(values = "orange") +
  facet_wrap(~survey_area, ncol = 1) +
  scale_y_continuous(labels = label_number(suffix = "M")) +
  labs(y = "Total Landings (live lb.)", x = "Year", color = "Moving Average")
```

### Zooplankton Community Indices



---


## Driver Correlations



## Driver Modeling

This section is likely to change

The following model has been used to determine the relationship between landings and seasonal bottom temperature on `b`.


```
lmer(
  b ~ survey_area*season*scale(roll_temp) + survey_area * scale(log(total_weight_lb)) + (1|est_year), 
  data = wtb_model_df)
```


```{r}

#### Model 2: Temperature & Fishing ####


# Drop NA's
wtb_model_df <- drop_na(wigley_bmspectra_model_df, total_weight_lb, bot_temp, b) %>%
  rename(area = survey_area) %>% 
  left_join(area_df) %>% 
  # Do rolling BT within a season * region
  group_by(survey_area, season) %>%
  mutate(
    roll_temp = zoo::rollapply(bot_temp, 5, mean, na.rm = T, align = "right",  fill = NA)) %>% 
  ungroup() %>% 
  mutate(
    yr_num = as.numeric(est_year),
    yr_fac = factor(est_year),
    survey_area = factor(survey_area, levels = area_levels),
    season = factor(season, levels = c("Spring", "Fall")),
    landings = total_weight_lb,
    yr_seas = str_c(season, est_year))


# Make the model
mod2 <- lmer(
  b ~ survey_area*season*scale(roll_temp) + survey_area*scale(log(total_weight_lb)) + (1|est_year), 
  data = wtb_model_df)


# summary(mod2)
# check_model(mod2)



```



### Temperature Relationship


```{r}
#| label: new bt plot


# Clean up the plot:
# Plot the predictions over data
mod2_preds <- as.data.frame(
  ggpredict(
    mod2, 
    terms = ~ roll_temp*survey_area*season))   %>% 
  mutate(
    survey_area = factor(group, levels = area_levels),
    season = factor(facet, levels = c("Spring", "Fall")))



#### Trend Posthoc  ####
trend_df <- emtrends(
  object = mod2,
  ~survey_area * season,
  var = "roll_temp")


# Drop effect fits that are non-significant  ###
mod2_emtrend <- emtrends(
  object = mod2,
  specs =  ~ survey_area*season,
  var = "roll_temp") %>%
  as_tibble() %>%
  mutate(
    zero = 0,
    non_zero = if_else(between(zero, lower.CL, upper.CL), F, T))




# Limit temperature plotting range
actual_range <- wtb_model_df %>% 
  group_by(season, survey_area) %>% 
  summarise(min_temp = min(bot_temp)-2,
            max_temp = max(bot_temp)+2,
            .groups = "drop")



# Plot over observed data
mod2_preds %>% 
  left_join(actual_range) %>% 
  filter((x < min_temp) == F,
         (x > max_temp) == F) %>% 
  left_join(select(mod2_emtrend, survey_area, season, non_zero)) %>%
  filter(non_zero) %>%
  ggplot() +
  geom_ribbon(aes(x, ymin = conf.low, ymax = conf.high, group = survey_area), alpha = 0.1) +
  geom_line(
    aes(x, predicted, color = season), 
    linewidth = 1) +
  geom_point(
    data = wtb_model_df,
    aes(roll_temp, b, color = season),
    alpha = 0.4,
    size = 1) +
  facet_grid(survey_area~., scales = "free") +
  scale_color_gmri() +
  scale_x_continuous(labels = label_number(suffix = deg_c)) +
  labs(y = "Exponent of ISD",
       title = "Wigley Species, Body Mass Spectra",
       x = "5-Year Rolling Average Bottom Temperature")

```

### Landings Relationship

```{r}
#| label: new landings

# Clean up the plot:
# Plot the predictions over data
mod2_preds <- as.data.frame(
  ggpredict(
    mod2, 
    terms = ~ total_weight_lb * survey_area * season))   %>% 
  mutate(
    survey_area = factor(group, levels = area_levels),
    season = factor(facet, levels = c("Spring", "Fall")))



#### Trend Posthoc  ####
trend_df <- emtrends(
  object = mod2,
  ~survey_area * season,
  var = "total_weight_lb")
#summary(trend_df, infer= c(T,T))


# Drop effect fits that are non-significant  ###
mod2_emtrend <- emtrends(
  object = mod2,
  specs =  ~ survey_area*season,
  var = "log(total_weight_lb)") %>%
  as_tibble() %>%
  mutate(
    zero = 0,
    non_zero = if_else(between(zero, lower.CL, upper.CL), F, T),
    group = str_c(survey_area, "X", season))




# Plot over observed data
mod2_preds %>% 
  left_join(select(mod2_emtrend, survey_area, season, non_zero)) %>%
  filter(non_zero) %>%
  ggplot() +
  geom_ribbon(aes(x, ymin = conf.low, ymax = conf.high, group = season), alpha = 0.1) +
  geom_line(
    aes(x, predicted, color = season), 
    linewidth = 1) +
  geom_point(
    data = wtb_model_df,
    aes(total_weight_lb, b, color = season),
    alpha = 0.4,
    size = 1) +
  facet_grid(survey_area~., scales = "free") +
  scale_color_gmri() +
  scale_x_continuous(transform = "log10", labels = label_log(base = 10)) +
  labs(y = "Body Mass Spectra Slope (b)",
       title = "Wigley Species, Body Mass Spectra",
       x = "Total Landings (lb.)")


```




### Driver Model Summary



```{r}
# mod2 %>% 
#   gtsummary::tbl_regression()   %>% 
#   add_global_p(keep = T) %>% 
#   bold_p(t = 0.05)  %>%
#   bold_labels() %>%
#   italicize_levels() %>% 
#   as_gt() %>% 
#   tab_header(title = "Wigley Community - Bodymass Distribution & Drivers")
```





## Side Stories

Everything below is digging into different nuances of the approach or the community composition. These things are of secondary importance and will be moved out.

### Spiny Dogfish Sensitivity

Spiny dogfish constitute a large majority fraction of biomass in the MAB during the Spring survey. Their numbers have also been increasing over time. This acts to shallow the spectra as they are one of the larger species routinely sampled by the trawl survey.

If we remove them from the estimation and re-run the slopes this is what changes. 

How much can one species shift the spectra, and is this a definitive/conclusive proof that its the dogfish shallowing the slope in MAB?

```{r}
#| eval: true
#| label: spiny dogfish spectra



# # Run MAB with and without spiny dogfish
# # Run the year, season, area fits for the filtered data
# no_dogs_bmspectra <- group_binspecies_spectra(
#   ss_input = filter(wigley_in, comname != "spiny dogfish"),
#   grouping_vars = c("est_year", "season", "survey_area"),
#   abundance_vals = "numlen_adj",
#   length_vals = "length_cm",
#   use_weight = TRUE,
#   isd_xmin = 1,
#   global_min = TRUE,
#   isd_xmax = NULL,
#   global_max = FALSE,
#   bin_width = 1,
#   vdiff = 2)

# # Save those
# write_csv(
#   no_dogs_bmspectra,
#   here::here("Data/processed/wigley_species_nodogfish_bmspectra.csv"))
```


```{r}
#| eval: true

# Read them in, do some plotting
no_dogs_bmspectra <- read_csv(
  here::here("Data/processed/wigley_species_nodogfish_bmspectra.csv")) %>% 
  mutate(survey_area = factor(survey_area, levels = area_levels))


# Format a little
no_dogs_bmspectra <- no_dogs_bmspectra %>% 
  mutate(est_year = as.numeric(est_year))


# Join them together
dfish_results <- bind_rows(
   list(
     "Wigley Full" = wigley_bmspectra,
     "Wigley w/o Spiny Dogfish" =  no_dogs_bmspectra), 
   .id = "community") %>% 
  mutate(metric = "bodymass_spectra") %>% 
  mutate(
    survey_area = fct_relevel(survey_area, area_levels))




# Get trends for no dogfish
nodfish_mod <- lmer(
  formula = b ~ est_year * survey_area * season + (1|est_year),
  data = no_dogs_bmspectra)
bmspectra_mod_wig <- lmer(
  formula = b ~ est_year * survey_area * season + (1|est_year),
  data = wigley_bmspectra)

# Put predictions
dogfish_sens_predictions <- bind_rows(list(
  "bodymass_spectra-Wigley Full" = get_preds_and_trendsignif(bmspectra_mod_wig),
  "bodymass_spectra-Wigley w/o Spiny Dogfish" = get_preds_and_trendsignif(nodfish_mod)), 
  .id = "model_id") %>% 
  separate(model_id, into = c("metric", "community"), sep = "-") %>% 
  mutate(metric = fct_rev(metric))



# Plot the differences
ggplot() +
  geom_ribbon(
    data = filter(dogfish_sens_predictions, non_zero == TRUE),
    aes(x, ymin = conf.low, ymax = conf.high, fill = season),
    linewidth = 1, alpha = 0.35) +
  geom_point(
    data = dfish_results,
    aes(est_year, b, color = season),
    size = 0.8, alpha = 0.5) +
  geom_line(
    data = filter(dogfish_sens_predictions, non_zero == TRUE),
    aes(x, predicted, color = season),
    linewidth = 1, alpha = 1) +
  scale_fill_gmri() +
  scale_color_gmri() +
  facet_grid(survey_area ~ metric*community, scales = "free") +
  labs(
    x = "Year",
    y = "Exponent of Size Spectra",
    title = "Shift in Spectra Trend(s) with Exclusion of Spiny Dogfish")


```




### Story Thoughts

Below is a table of some papers operating in the same area or around the same topic that are particularly relevant:

```{r}
shelf_papers <- tribble(
  ~"Author", ~"Year", ~"Conjecture",
  "Friedland et al.", 2024, "Body size has declined despite more abundance & biomass. Competition is preventing larger sizes.",
  "Krumsick", 2020, "Empirical slopes from trawl survey in labrador sea were steeper than theoretical from nitrogen stable isotope ratios"
)


shelf_papers %>% gt()
```


There are a couple core ideas about what changes have occurred or have been ocurring over the Northeast Shelf with relevance to the community that is sampled by the trawl survey:


__________________




